---
title: "covid 19"
author: "Amy Estrada"
date: "3/2/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Covid 19 deaths in Korea, January-May 2020 are unrelated to sex.
$H_A$: There is a relation between deaths in Korea January-May 2020 and sex of the patiant.

Table of patients:
```{r}
kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result = as.factor(kRdata$result)

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

This table shows how many males and females have survived COVID-19 and have passed away from COVID-19 in Korea.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result, Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The graph shows the results and counts of the males and females that survived and passed away from COVID-19 in Korea.

## Numerical results

```{r}
table1 <- xtabs(~Sex + result, kRdata)
rowPerc(table1)
colPerc(table1)
```

The table above shows the percentage of the males and females that have been affected by COVID-19, also shows you the count of who has passed away and survived COVID-19 in Korea.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This table demonstrates the observed counts, observed counts null, and the chi-square of the results for both hypothesis. The p-vale is to small to show that gender is a factor to COVID-19 deaths in Korea.

## Conclusions

In conclusion, we reject the alternative hypothesis. We accept our hypothesis; sex is not related to COVID-19 deaths in Korea. The p-value is to small to indicate that the deaths are related to sex.
